Sparse spectral and $p$-finite element methods for partial differential equations on disk slices and trapeziums

### 2020-01-06 papers/disk-slices-and-trapeziums/revised/halfdisk.tex
Further edits to paper, inc figs. Made some comments.

\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma} 
\newtheorem{theorem}{Theorem} 
\newtheorem{definition}{Definition}
\newtheorem{condition}{Condition}
\newtheorem{algorithm}{Algorithm}

@@

Sparsity comes from expanding the domain and range of an operator  using different choices of $a$ and $b$. Whereas the sparsity pattern and entries derived in \cite{olver2018recurrence,olver2019triangle} for equations on the triangle  and \cite{vasil2016tensor} for equations on the disk results from manipulations of Jacobi polynomials, in the present work we use a more general integration-by-parts argument to deduce the sparsity structure, alongside quadrature rules\sotodo{Mention Christoffel--Darboux here?} to determine the entries.  In particular, by exploiting the connection with one-dimensional orthogonal polynomials \bstodo{typo 2}we can construct discretizations of general partial differential operators of size $p(p-1)/2 \times p(p-1)/2$ in $O(p^3)$ operations, where $p$ is the total polynomial degree. This compares favourably to $O(p^6)$ operations if one proceeds na\"ively. Furthermore, we use this framework to derive sparse $p$-finite element methods that are analogous to those of Beuchler and Sch\"oberl on tetrahedra \cite{beuchler2006new}, see also work by Li and Shen \cite{li2010optimal}.

@@

\noindent\secref{PDOs}: We demonstrate that these families will lead to sparse operators, including Jacobi operators representing multiplication by $x$ and $y$, and partial differential operators.

\sotodoinline{Mention Christoffell--Darboux here}

@@

\begin{figure}[t]
	\begin{subfigure}{0.32\textwidth}
	\includegraphics[scale=0.35]{sparsityoflaplacian-w11} %-diskslice-alpha=0p2-beta=0p8}
        \centering
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
	\includegraphics[scale=0.35]{sparsityofhelmholtz} %-diskslice-alpha=0p2-beta=0p8}
        \centering
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
	\includegraphics[scale=0.35]{sparsityofbiharmonic} %-diskslice-alpha=0p2-beta=0p8}
        \centering
	\end{subfigure}
    	\caption{"Spy" plots of (differential) operator matrices, showing their sparsity. Left: the Laplace operator $\laplacewiii$. Centre: the weighted variable coefficient Helmholtz operator $\laplacewiii + k^2 \: T^{(0,0,0)\to(1,1,1)} \: V({J_x^{(0,0,0)}}^\top, {J_y^{(0,0,0)}}^\top) \: T_W^{(1,1,1)\to(0,0,0)}$ for $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 200$. Right: the biharmonic operator $\biharmonictwo$.}
        \label{fig:sparsity}
        \centering
\end{figure}

@@

It is possible to obtain the recurrence coefficients for the $\{\genjac_n^{(a,b,c)}\}$ OPs in (\ref{eqn:Hrecurrence}), by careful application of the Christoffel-Darboux formula \bstodo{Link!!!}. We explain the process here for the disk-slice case, however we note that a similar but simpler argument holds for the trapezium case. We thus first need to define a new set of `interim' 1D OPs.
\begin{definition}\label{def:InterimOPconstruction}
Let $\genjactw^{(a,b,c,d)}(x) := (\beta - x)^a \: (x - \alpha)^{b} \: (1-x)^{c} \: (1+x)^{d} $ be a weight function on the interval $(\alpha, \beta)$, and define the associated inner product by:
\begin{align}
	\ip< p, \: q >_{\genjactw^{(a,b,c,d)}} &:= \frac{1}{\normgenjact^{(a,b,c,d)}} \: \int_\alpha^\beta p(x) \: q(x) \: \genjactw^{(a,b,c,d)}(x) \: \D x \label{eqn:ipgenjact}
\end{align}
where
\begin{align}
	\normgenjact^{(a,b,c,d)} := \int_\alpha^\beta \: \genjactw^{(a,b,c,d)}(x) \: \D x
\end{align}
Denote the four-parameter family of orthonormal polynomials on $[\alpha,\beta]$ by $\{\genjact_n^{(a,b,c,d)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipgenjact}).
\end{definition}
Note that the OPs $\{\genjac_n^{(a,b,2c)}\}$ are then equivalent to the OPs $\{\genjact_n^{(a,b,c,c)}\}$. Let the recurrence coefficients for the OPs $\{\genjact_n^{(a,b,c,d)}\}$ be given by:
\begin{align}
	x \: \genjact_n^{(a,b,c,d)}(x) = \tilde{\beta}_n^{(a,b,c,d)} \: \genjact_{n+1}^{(a,b,c,d)}(x) + \tilde{\alpha}_n^{(a,b,c,d)} \:\genjact_n^{(a,b,c,d)}(x) + \tilde{\beta}_{n-1}^{(a,b,c,d)} \: \genjact_{n-1}^{(a,b,c,d)}(x)
\end{align}
\begin{proposition}
There exist constants $\mathcal{C}_n^{(a,b,c,d)}$, $\mathcal{D}_n^{(a,b,c,d)}$ such that
\begin{align}
	\genjact_n^{(a,b,c+1,d)}(x) &= \mathcal{C}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x) \label{eqn:genjacexpansioni} \\
	\genjact_n^{(a,b,c,d+1)}(x) &= \mathcal{D}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x) \label{eqn:genjacexpansionii}
\end{align}
\end{proposition}

\begin{proof}
Fix $n,m \in \{0,1,\dots\}$ and without loss of generality, assume $m \le n$. First recall that
\begin{align*}
	\int_\alpha^\beta \: \genjact_n^{(a,b,c+1,d)}(x) \: \genjact_m^{(a,b,c+1,d)}(x) \: \genjactw^{(a,b,c+1,d)}(x) \: \D x &= \delta_{n,m} \: \normgenjact^{(a,b,c+1,d)}
\end{align*}
and define \bstodo{check signs}
\begin{align}
	\mathcal{C}_n^{(a,b,c,d)} &= \fpr(\frac{\normgenjact^{(a,b,c+1,d)}}{\normgenjact^{(a,b,c,d)} \: \genjact_n^{(a,b,c,d)}(1) \: \genjact_{n+1}^{(a,b,c,d)}(1) \: \tilde{\beta}_n^{(a,b,c,d)}})^\half, \label{eqn:genjactnormalisationi} \\
	\mathcal{D}_n^{(a,b,c,d)} &= (-1)^n \: \fpr(\frac{- \normgenjact^{(a,b,c,d+1)}}{\normgenjact^{(a,b,c,d)} \: \genjact_n^{(a,b,c,d)}(-1) \: \genjact_{n+1}^{(a,b,c,d)}(-1) \: \tilde{\beta}_n^{(a,b,c,d)}})^\half. \label{eqn:genjactnormalisationii}
\end{align}
Now, by the Christoffel-Darboux formula \bstodo{cite}, we have that for any $x, y \in \R$,
\begin{align}
	\sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(y) \: \genjact_{k}^{(a,b,c,d)}(x) &= \tilde{\beta}_n^{(a,b,c,d)} \: \frac{\genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(y) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(y)}{y - x}. \label{eqn:christoffeldarboux}
\end{align}
Then,
\begin{align*}
	&\int_\alpha^\beta \: \Big(\big[\mathcal{C}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \\
	&\quad \quad \quad \cdot \big[\mathcal{C}_m^{(a,b,c,d)} \: \sum_{k=0}^m \: \genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \: \genjactw^{(a,b,c+1,d)}(x) \Big) \: \D x \\
	&= \mathcal{C}_n^{(a,b,c,d)} \: \mathcal{C}_m^{(a,b,c,d)} \: \tilde{\beta}_n^{(a,b,c,d)} \\
	&\quad \quad \cdot \sum_{k=0}^m \: \int_\alpha^\beta \: \Big(\genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x) \\
	&\quad \quad \quad \quad \quad \quad \quad \cdot \big[\genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(1) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(1)\big] \: \genjactw^{(a,b,c,d)}(x)\Big) \: \D x \\
	&= \delta_{m,n} \: {\mathcal{C}_n^{(a,b,c,d)}}^2 \: \tilde{\beta}_n^{(a,b,c,d)} \: \normgenjact^{(a,b,c,d)} \: \genjact_{n}^{(a,b,c,d)}(1) \: \genjact_{n+1}^{(a,b,c,d)}(1) \\
	&= \delta_{m,n} \: \normgenjact^{(a,b,c+1,d)}
\end{align*}
using (\ref{eqn:genjactnormalisationi}) and (\ref{eqn:christoffeldarboux}), showing that the RHS and LHS of (\ref{eqn:genjacexpansioni}) are equivalent. Further,
\begin{align*}
	&\int_\alpha^\beta \: \Big(\big[\mathcal{D}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \\
	&\quad \quad \quad \cdot \big[\mathcal{D}_m^{(a,b,c,d)} \: \sum_{k=0}^m \: \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \: \genjactw^{(a,b,c,d+1)}(x) \Big) \: \D x \\
	&= - \mathcal{D}_n^{(a,b,c,d)} \: \mathcal{D}_m^{(a,b,c,d)} \: \tilde{\beta}_n^{(a,b,c,d)} \\
	&\quad \quad \cdot \sum_{k=0}^m \: \int_\alpha^\beta \: \Big(\genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x) \\
	&\quad \quad \quad \quad \quad \quad \quad \cdot \big[\genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(-1) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(-1)\big] \: \genjactw^{(a,b,c,d)}(x)\Big) \: \D x \\
	&= - \delta_{m,n} \: {\mathcal{D}_n^{(a,b,c,d)}}^2 \: \tilde{\beta}_n^{(a,b,c,d)} \: \normgenjact^{(a,b,c,d)} \: \genjact_{n}^{(a,b,c,d)}(-1) \: \genjact_{n+1}^{(a,b,c,d)}(-1) \\
	&= \delta_{m,n} \: \normgenjact^{(a,b,c,d+1)}
\end{align*}
using (\ref{eqn:genjactnormalisationii}) and (\ref{eqn:christoffeldarboux}), showing that the RHS and LHS of (\ref{eqn:genjacexpansionii}) are also equivalent. 
\end{proof}

@@

@@

\noindent{\bf Remark}:  \bstodo{improve this remark}
This is the most expensive part of the current calculation, in particular the initialisation required (i.e. using the Lanczos algorithm to compute the recurrence coefficients $\{\alpha_{n}^{(a,b,2c+1)}\}$, $\{\beta_{n}^{(a,b,2c+1)}\}$ -- however, we only need to compute these once, and can store and save this initialisation to disk once computed, for the given values of $a, b, c$). Note also that in order to obtain the recurrence coefficients $\{\alpha_{n-k}^{(a,b,2c+2k+1)}\}$, $\{\beta_{n-k}^{(a,b,2c+2k+1)}\}$ for $n = 0,\dots,N$ for some $N$ and $k$, we require that we obtain the recurrence coefficients $\{\alpha_{n}^{(a,b,2c+2(k-1)+1)}\}$, $\{\beta_{n}^{(a,b,2c+2(k-1)+1)}\}$ for $n = 0,\dots,N+3$. Thus, for large $k$, this recursive method of obtaining the recurrence coefficients requires a large initialisation.
%This is the most expensive part of the current calculation, but note that we can reuse this computation for multiple partial differential equations. Efficient construction of OPs with large parameters in the weights is an important topic, but one tangential to the proposed scheme.

@@

% FIGURES
\begin{figure}[t]
	\begin{subfigure}{0.3\textwidth}
	\includegraphics[scale=0.3]{solution-poisson}
	\centering
	%\label{fig:solution-poisson}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
	\centering
	\includegraphics[scale=0.5]{solutionblocknorms-poisson} %-diskslice-alpha=0p2-beta=0p8}
        	%\label{fig:solutionblocknorms-poisson}
	\end{subfigure}
	\caption{Left: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$. Right: The norms of each block of the computed solution of the Poisson equation with the given right hand side functions. This demonstrates algebraic convergence with the rate dictated by the decay at the corners, with spectral convergence observed when the right-hand side vanishes to all orders.}
	\centering
	\label{fig:poisson}
\end{figure}

\begin{figure}[t]
	\begin{subfigure}{0.3\textwidth}
	\centering
	\includegraphics[scale=0.3]{solution-helmholtz-k=100-n=1500}
	%\label{fig:solution-helmholtz}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
	\includegraphics[scale=0.5]{solutionblocknorms-helmholtz-k=20} %-diskslice-alpha=0p2-beta=0p8-N=196-k=20}
	\centering
        	%\label{fig:solutionblocknorms-helmholtz}
	\end{subfigure}
	\caption{Left: The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y) = x(1-x^2-y^2)e^x$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 100$. Right: The norms of each block of the computed solution of the Helmholtz equation with the given right hand side functions, with $k=20$ and $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$.}
	\centering
	\label{fig:helmholtz}
\end{figure}

\begin{figure}[t]
	\begin{subfigure}{0.3\textwidth}
	\centering
	\includegraphics[scale=0.3]{solution-biharmonic}
	%\label{fig:solution-biharmonic}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
	\includegraphics[scale=0.5]{solutionblocknorms-biharmonic}
	\centering
        	%\label{fig:solutionblocknorms-biharmonic}
	\end{subfigure}
	\caption{Left: The computed solution to $\Delta^2 u = f$ with zero Dirichlet and Neumann boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$. Right: The norms of each block of the computed solution of the biharmonic equation with the given right hand side functions.}
	\centering
	\label{fig:biharmonic}
\end{figure}
% END FIGURES

@@

@@

>>>>

\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma} 
\newtheorem{theorem}{Theorem} 
\newtheorem{definition}{Definition}
\newtheorem{condition}{Condition}
\newtheorem{algorithm}{Algorithm}
\newtheorem{corollary}{Corollary}

@@

Sparsity comes from expanding the domain and range of an operator using different choices of $a$ and $b$. Whereas the sparsity pattern and entries derived in \cite{olver2018recurrence,olver2019triangle} for equations on the triangle and \cite{vasil2016tensor} for equations on the disk results from manipulations of Jacobi polynomials, in the present work we use a more general integration-by-parts argument to deduce the sparsity structure, alongside careful use of the Christoffel-Darboux formula \cite[18.2.2]{DLMF} and quadrature rules to determine the entries. In particular, by exploiting the connection with one-dimensional orthogonal polynomials \bstodo{typo 2}we can construct discretizations of general partial differential operators of size $p(p-1)/2 \times p(p-1)/2$ in $O(p^3)$ operations, where $p$ is the total polynomial degree. This compares favourably to $O(p^6)$ operations if one proceeds na\"ively. Furthermore, we use this framework to derive sparse $p$-finite element methods that are analogous to those of Beuchler and Sch\"oberl on tetrahedra \cite{beuchler2006new}, see also work by Li and Shen \cite{li2010optimal}.

@@

\noindent\secref{PDOs}: We demonstrate that these families will lead to sparse operators, including Jacobi operators representing multiplication by $x$ and $y$, and partial differential operators. We present a method involving use of the Christoffel-Darboux formula \cite[18.2.2]{DLMF} to obtain the recurrence coefficients for the non-classical 1D OPs, allowing us to exactly use the 1D quadrature rules we present to calculate the non-zero entries to the sparse operators.

@@

\begin{figure}[t]
	\begin{subfigure}{0.32\textwidth}
	\includegraphics[scale=0.35]{sparsityoflaplacian-w11-diskslice-alpha=0p2-beta=0p8}
        \centering
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
	\includegraphics[scale=0.35]{sparsityofhelmholtz-diskslice-alpha=0p2-beta=0p8}
        \centering
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
	\includegraphics[scale=0.35]{sparsityofbiharmonic-diskslice-alpha=0p2-beta=0p8}
        \centering
	\end{subfigure}
    	\caption{"Spy" plots of (differential) operator matrices, showing their sparsity. Left: the Laplace operator $\laplacewiii$. Centre: the weighted variable coefficient Helmholtz operator $\laplacewiii + k^2 \: T^{(0,0,0)\to(1,1,1)} \: V({J_x^{(0,0,0)}}^\top, {J_y^{(0,0,0)}}^\top) \: T_W^{(1,1,1)\to(0,0,0)}$ for $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 200$. Right: the biharmonic operator $\biharmonictwo$.}
        \label{fig:sparsity}
        \centering
\end{figure}

@@

It is possible to obtain the recurrence coefficients for the $\{\genjac_n^{(a,b,c)}\}$ OPs in (\ref{eqn:Hrecurrence}), by careful application of the Christoffel-Darboux formula \cite[18.2.12]{DLMF}. We explain the process here for the disk-slice case, however we note that a similar but simpler argument holds for the trapezium case. We thus first need to define a new set of `interim' 1D OPs.
\begin{definition}\label{def:InterimOPconstruction}
Let $\genjactw^{(a,b,c,d)}(x) := (\beta - x)^a \: (x - \alpha)^{b} \: (1-x)^{c} \: (1+x)^{d} $ be a weight function on the interval $(\alpha, \beta)$, and define the associated inner product by:
\begin{align}
	\ip< p, \: q >_{\genjactw^{(a,b,c,d)}} &:= \frac{1}{\normgenjact^{(a,b,c,d)}} \: \int_\alpha^\beta p(x) \: q(x) \: \genjactw^{(a,b,c,d)}(x) \: \D x \label{eqn:ipgenjact}
\end{align}
where
\begin{align}
	\normgenjact^{(a,b,c,d)} := \int_\alpha^\beta \: \genjactw^{(a,b,c,d)}(x) \: \D x
\end{align}
Denote the four-parameter family of orthonormal polynomials on $[\alpha,\beta]$ by $\{\genjact_n^{(a,b,c,d)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipgenjact}).
\end{definition}
Note that the OPs $\{\genjac_n^{(a,b,2c)}\}$ are then equivalent to the OPs $\{\genjact_n^{(a,b,c,c)}\}$. Let the recurrence coefficients for the OPs $\{\genjact_n^{(a,b,c,d)}\}$ be given by:
\begin{align}
	x \: \genjact_n^{(a,b,c,d)}(x) = \tilde{\beta}_n^{(a,b,c,d)} \: \genjact_{n+1}^{(a,b,c,d)}(x) + \tilde{\alpha}_n^{(a,b,c,d)} \:\genjact_n^{(a,b,c,d)}(x) + \tilde{\beta}_{n-1}^{(a,b,c,d)} \: \genjact_{n-1}^{(a,b,c,d)}(x)
\end{align}
\begin{proposition}
There exist constants $\mathcal{C}_n^{(a,b,c,d)}$, $\mathcal{D}_n^{(a,b,c,d)}$ such that
\begin{align}
	\genjact_n^{(a,b,c+1,d)}(x) &= \mathcal{C}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x) \label{eqn:genjacexpansioni} \\
	\genjact_n^{(a,b,c,d+1)}(x) &= \mathcal{D}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x) \label{eqn:genjacexpansionii}
\end{align}
\end{proposition}

\begin{proof}
Fix $n,m \in \{0,1,\dots\}$ and without loss of generality, assume $m \le n$. First recall that
\begin{align*}
	\int_\alpha^\beta \: \genjact_n^{(a,b,c+1,d)}(x) \: \genjact_m^{(a,b,c+1,d)}(x) \: \genjactw^{(a,b,c+1,d)}(x) \: \D x &= \delta_{n,m} \: \normgenjact^{(a,b,c+1,d)}
\end{align*}
and define
\begin{align}
	\mathcal{C}_n^{(a,b,c,d)} &= \fpr(\frac{\normgenjact^{(a,b,c+1,d)}}{\normgenjact^{(a,b,c,d)} \: \genjact_n^{(a,b,c,d)}(1) \: \genjact_{n+1}^{(a,b,c,d)}(1) \: \tilde{\beta}_n^{(a,b,c,d)}})^\half, \label{eqn:genjactnormalisationi} \\
	\mathcal{D}_n^{(a,b,c,d)} &= (-1)^n \: \fpr(\frac{- \normgenjact^{(a,b,c,d+1)}}{\normgenjact^{(a,b,c,d)} \: \genjact_n^{(a,b,c,d)}(-1) \: \genjact_{n+1}^{(a,b,c,d)}(-1) \: \tilde{\beta}_n^{(a,b,c,d)}})^\half. \label{eqn:genjactnormalisationii}
\end{align}
Now, by the Christoffel--Darboux formula \cite[18.2.12]{DLMF}, we have that for any $x, y \in \R$,
\begin{align}
	\sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(y) \: \genjact_{k}^{(a,b,c,d)}(x) &= \tilde{\beta}_n^{(a,b,c,d)} \: \frac{\genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(y) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(y)}{y - x}. \label{eqn:christoffeldarboux}
\end{align}
Then,
\begin{align*}
	&\int_\alpha^\beta \: \Big(\big[\mathcal{C}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \\
	&\quad \quad \quad \cdot \big[\mathcal{C}_m^{(a,b,c,d)} \: \sum_{k=0}^m \: \genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \: \genjactw^{(a,b,c+1,d)}(x) \Big) \: \D x \\
	&= \mathcal{C}_n^{(a,b,c,d)} \: \mathcal{C}_m^{(a,b,c,d)} \: \tilde{\beta}_n^{(a,b,c,d)} \\
	&\quad \quad \cdot \sum_{k=0}^m \: \int_\alpha^\beta \: \Big(\genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x) \\
	&\quad \quad \quad \quad \quad \quad \quad \cdot \big[\genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(1) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(1)\big] \: \genjactw^{(a,b,c,d)}(x)\Big) \: \D x \\
	&= \delta_{m,n} \: {\mathcal{C}_n^{(a,b,c,d)}}^2 \: \tilde{\beta}_n^{(a,b,c,d)} \: \normgenjact^{(a,b,c,d)} \: \genjact_{n}^{(a,b,c,d)}(1) \: \genjact_{n+1}^{(a,b,c,d)}(1) \\
	&= \delta_{m,n} \: \normgenjact^{(a,b,c+1,d)}
\end{align*}
using (\ref{eqn:genjactnormalisationi}) and (\ref{eqn:christoffeldarboux}), showing that the RHS and LHS of (\ref{eqn:genjacexpansioni}) are equivalent. Further,
\begin{align*}
	&\int_\alpha^\beta \: \Big(\big[\mathcal{D}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \\
	&\quad \quad \quad \cdot \big[\mathcal{D}_m^{(a,b,c,d)} \: \sum_{k=0}^m \: \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \: \genjactw^{(a,b,c,d+1)}(x) \Big) \: \D x \\
	&= - \mathcal{D}_n^{(a,b,c,d)} \: \mathcal{D}_m^{(a,b,c,d)} \: \tilde{\beta}_n^{(a,b,c,d)} \\
	&\quad \quad \cdot \sum_{k=0}^m \: \int_\alpha^\beta \: \Big(\genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x) \\
	&\quad \quad \quad \quad \quad \quad \quad \cdot \big[\genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(-1) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(-1)\big] \: \genjactw^{(a,b,c,d)}(x)\Big) \: \D x \\
	&= - \delta_{m,n} \: {\mathcal{D}_n^{(a,b,c,d)}}^2 \: \tilde{\beta}_n^{(a,b,c,d)} \: \normgenjact^{(a,b,c,d)} \: \genjact_{n}^{(a,b,c,d)}(-1) \: \genjact_{n+1}^{(a,b,c,d)}(-1) \\
	&= \delta_{m,n} \: \normgenjact^{(a,b,c,d+1)}
\end{align*}
using (\ref{eqn:genjactnormalisationii}) and (\ref{eqn:christoffeldarboux}), showing that the RHS and LHS of (\ref{eqn:genjacexpansionii}) are also equivalent. 
\end{proof}

@@

\begin{corollary}
The recurrence coefficients for the OPs $\{\genjact_n^{(a,b,c+1,d)}\}$ can be written as:
\begin{align}
	\tilde{\alpha}_n^{(a,b,c+1,d)} &= \frac{\tilde{\beta}_{n-1}^{(a,b,c,d)}}{\chi_{n-1}^{(a,b,c,d)}(1)} - \frac{\tilde{\beta}_{n}^{(a,b,c,d)}}{\chi_{n}^{(a,b,c,d)}(1)} + \tilde{\alpha}_{n}^{(a,b,c,d)}, \\
	\tilde{\beta}_n^{(a,b,c+1,d)} &= \fpr(\frac{1 - \tilde{\alpha}_{n+1}^{(a,b,c,d)} - \frac{\tilde{\beta}_n^{(a,b,c,d)}}{\chi_n^{(a,b,c,d)}(1)}}{1 - \tilde{\alpha}_{n}^{(a,b,c,d)} - \frac{\tilde{\beta}_{n-1}^{(a,b,c,d)}}{\chi_{n-1}^{(a,b,c,d)}(1)}})^\half \tilde{\beta}_n^{(a,b,c,d)}.
\end{align}
The recurrence coefficients for the OPs $\{\genjact_n^{(a,b,c,d+1)}\}$ can be written as:
\begin{align}
	\tilde{\alpha}_n^{(a,b,c,d+1)} &= \frac{\tilde{\beta}_{n-1}^{(a,b,c,d)}}{\chi_{n-1}^{(a,b,c,d)}(-1)} - \frac{\tilde{\beta}_{n}^{(a,b,c,d)}}{\chi_{n}^{(a,b,c,d)}(-1)} + \tilde{\alpha}_{n}^{(a,b,c,d)}, \\
	\tilde{\beta}_n^{(a,b,c,d+1)} &= \fpr(\frac{- 1 + \tilde{\alpha}_{n+1}^{(a,b,c,d)} + \frac{\tilde{\beta}_n^{(a,b,c,d)}}{\chi_n^{(a,b,c,d)}(-1)}}{- 1 + \tilde{\alpha}_{n}^{(a,b,c,d)} + \frac{\tilde{\beta}_{n-1}^{(a,b,c,d)}}{\chi_{n-1}^{(a,b,c,d)}(-1)}})^\half \tilde{\beta}_n^{(a,b,c,d)}.
\end{align}
where 
\begin{align}
	\chi_{n}^{(a,b,c,d)}(y) &:= \frac{\genjact_{n+1}^{(a,b,c,d)}(y)}{\genjact_{n}^{(a,b,c,d)}(y)} \\
	&= \frac{1}{\tilde{\beta}_{n}^{(a,b,c,d)}} \: \fpr(y - \tilde{\alpha}_{n}^{(a,b,c,d)} - \frac{\tilde{\beta}_{n-1}^{(a,b,c,d)}}{\chi_{n-1}^{(a,b,c,d)}(y)}), \quad y \in \{-1,1\}.
\end{align}
\end{corollary}

@@

\noindent{\bf Remark}:
The Corollary demonstrates that in order to obtain the recurrence coefficients $\{\alpha_{m}^{(a,b,2c+2k+1)}\}$, $\{\beta_{m}^{(a,b,2c+2k+1)}\}$ for some $m$ and $k$, we require that we obtain the recurrence coefficients $\{\alpha_{m+2}^{(a,b,2c+2(k-1)+1)}\}$, $\{\beta_{m+2}^{(a,b,2c+2(k-1)+1)}\}$. Thus, for large $N$, this recursive method of obtaining the recurrence coefficients requires a large initialisation (i.e. using the Lanczos algorithm to compute the recurrence coefficients $\{\alpha_{n}^{(a,b,2c+1)}\}$, $\{\beta_{n}^{(a,b,2c+1)}\}$ -- however, we only need to compute these once, and can store and save this initialisation to disk once computed, for the given values of $a, b, c$). 
%This is the most expensive part of the current calculation, but note that we can reuse this computation for multiple partial differential equations. Efficient construction of OPs with large parameters in the weights is an important topic, but one tangential to the proposed scheme.

@@

% FIGURES
\begin{figure}[t]
	\begin{subfigure}{0.3\textwidth}
	\includegraphics[scale=0.3]{solution-poisson-diskslice-alpha=0p2-beta=0p8}
	\centering
	%\label{fig:solution-poisson}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
	\centering
	\includegraphics[scale=0.5]{solutionblocknorms-poisson-diskslice-alpha=0p2-beta=0p8-N=200}
        	%\label{fig:solutionblocknorms-poisson}
	\end{subfigure}
	\caption{Left: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$. Right: The norms of each block of the computed solution of the Poisson equation with the given right hand side functions. This demonstrates algebraic convergence with the rate dictated by the decay at the corners, with spectral convergence observed when the right-hand side vanishes to all orders.}
	\centering
	\label{fig:poisson}
\end{figure}

\begin{figure}[t]
	\begin{subfigure}{0.3\textwidth}
	\centering
	\includegraphics[scale=0.3]{solution-helmholtz-diskslice-alpha=0p2-beta=0p8-k=100-n=300}
	%\label{fig:solution-helmholtz}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
	\includegraphics[scale=0.5]{solutionblocknorms-helmholtz-diskslice-alpha=0p2-beta=0p8-N=196-k=20}
	\centering
        	%\label{fig:solutionblocknorms-helmholtz}
	\end{subfigure}
	\caption{Left: The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y) = x(1-x^2-y^2)e^x$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 100$. Right: The norms of each block of the computed solution of the Helmholtz equation with the given right hand side functions, with $k=20$ and $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$.}
	\centering
	\label{fig:helmholtz}
\end{figure}

\begin{figure}[t]
	\begin{subfigure}{0.3\textwidth}
	\centering
	\includegraphics[scale=0.3]{solution-biharmonic-diskslice-alpha=0p2-beta=0p8}
	%\label{fig:solution-biharmonic}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
	\includegraphics[scale=0.5]{solutionblocknorms-biharmonic-diskslice-alpha=0p2-beta=0p8-N=200}
	\centering
        	%\label{fig:solutionblocknorms-biharmonic}
	\end{subfigure}
	\caption{Left: The computed solution to $\Delta^2 u = f$ with zero Dirichlet and Neumann boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$. Right: The norms of each block of the computed solution of the biharmonic equation with the given right hand side functions.}
	\centering
	\label{fig:biharmonic}
\end{figure}
% END FIGURES

@@

In Figure \ref{fig:appendixfigs} we see the solution to the Poisson equation with zero boundary conditions in the half-disk $\Omega$ with $(\alpha,\beta) := (0,1)$.

@@

In Figure \ref{fig:appendixfigs} we see the solution to the Helmholtz equation with zero boundary conditions in the trapezium $\Omega$ with $\xi := \half$.
